Word learning at the semantics pragmatics interface

### 2020-03-31 paper/supplementary_material.Rmd
work on MHs comments

The informativity of each utterance is given by imagining which referent a literal listener, who interprets words according to their literal semantics, would infer upon hearing the utterance. Thus, this reasoning depends on what kind of semantic knowledge (word--object mappings) the speaker thinks the literal listener knows. We assume that a listener's knowledge of a word's semantics is graded (not categorical). That is, for each object involved there is a probability $p$ that the literal listener knows the word for it. For each of the novel objects, this semantic knowledge is 0, which assumes that there is a no *a priori* reason to associate a particular novel object with a particular novel word. (MH: <-- is this what this represents? or does it represent the fact that there is no reason to associate a novel object with a familiar word?) For familiar objects, semantic knowledge is a function of the degree-of-acquisition of the associated word, which in turn depends upon the kind of object (its expected acquisition trajectory) as well as on the child's age.

@@

Semantic knowledge captures the degree of certainty with which the naive listener is assumed to know the label for the familiar object. (MH: UNCLEAR --> When faced with the task, we think that children take their own semantic knowledge as the basis.) As a consequence, semantic knowledge differs between familiar objects. For objects whose labels are generally acquired earlier (e.g. carrot) semantic knowledge should generally be high whereas for others (e.g. pawn) semantic knowledge should generally be lower. However, semantic knowledge also varies with age such that older children are more likely to know the labels for more of the familiar objects compared to younger children. As a consequence, each familiar object has a unique developmental trajectory with respect to semantic knowledge. Technically, the object-specific parameters are estimated in the form of a hierarchical regression (mixed-effects) model; that is, each object's trajectory is estimated as a deviation from an overall trajectory of vocabulary development. This overall (vocabulary) trajectory represents the development of semantic knowledge that is independent of a particular familiar object--label pairing.

@@

A second locus of developmental change is a listener's expectations about speaker informativeness. In the context of the model, speaker informativeness corresponds to the degree with which the listener expects the speaker to choose the most informative of the available utterances. We assume that children at different ages could have different expectations about how rational or informative speakers are. (MH: citation here, kids sensitivity to informativeness?) As mentioned above, this parameter is jointly estimated with the parameters for semantic knowledge.

@@

Sensitivity to common ground refers to the probability that an object is taken to be the referent of the utterance before actually hearing the utterance. Thus, it captures the salience of an object due to its role in the social interaction that precedes the utterance. We expect children at different ages to respond differently to the common ground manipulation, resulting in an age specific prior distribution over objects. (MH: again, citation?)

@@

All Bayesian cognitive models were implemented in the probabilistic programming language `WebPPL` [@dippl]. The corresponding model code can be found in the associated online repository and includes information about the prior distributions for all parameters (file xxxxx). To generate model predictions, we estimated age sensitive parameter distributions for semantic knowledge (by familiar object), speaker informativeness and common ground sensitivity and then passed them through the model in line with the different ways in which they can be combined and aligned. The resulting predictions come in the form of distributions of developmental trajectories for each object in the congruent and the incongruent condition.  (MH: Priors for parameters should be reported here.)

>>>>

The informativity of each utterance is given by imagining which referent a literal listener, who interprets words according to their literal semantics, would infer upon hearing the utterance. Thus, this reasoning depends on what kind of semantic knowledge (word--object mappings) the speaker thinks the literal listener knows. We assume that a listener's knowledge of a word's semantics is graded (not categorical). That is, for each object involved there is a probability $p$ that the literal listener knows the word for it. For each of the novel objects, this semantic knowledge is 0, which assumes that the novel word is equaliy likely to refer to the novel object as it is to refer to the familiar object. For familiar objects, semantic knowledge is a function of the degree-of-acquisition of the associated word, which in turn depends upon the kind of object (its expected acquisition trajectory) as well as on the child's age.

@@

Semantic knowledge captures the degree of certainty with which the naive listener is assumed to know the label for the familiar object. As a consequence, semantic knowledge differs between familiar objects. For objects whose labels are generally acquired earlier (e.g. carrot) semantic knowledge should generally be high whereas for others (e.g. pawn) semantic knowledge should generally be lower. However, semantic knowledge also varies with age such that older children are more likely to know the labels for more of the familiar objects compared to younger children. As a consequence, each familiar object has a unique developmental trajectory with respect to semantic knowledge. Technically, the object-specific parameters are estimated in the form of a hierarchical regression (mixed-effects) model; that is, each object's trajectory is estimated as a deviation from an overall trajectory of vocabulary development. This overall (vocabulary) trajectory represents the development of semantic knowledge that is independent of a particular familiar object--label pairing.

@@

A second locus of developmental change is a listener's expectations about speaker informativeness. In the context of the model, speaker informativeness corresponds to the degree with which the listener expects the speaker to choose the most informative of the available utterances. We assume that children at different ages could have different expectations about how rational or informative speakers are [see e.g. @bohn_tessler_merrick_frank_2019; @frank2014inferring; @yoon2019role]. As mentioned above, this parameter is jointly estimated with the parameters for semantic knowledge.

@@

Sensitivity to common ground refers to the probability that an object is taken to be the referent of the utterance before actually hearing the utterance. Thus, it captures the salience of an object due to its role in the social interaction that precedes the utterance. We expect children at different ages to respond differently to the common ground manipulation, resulting in an age specific prior distribution over objects [akhtar1996role; @diesendruck2004two].

@@

All Bayesian cognitive models were implemented in the probabilistic programming language `WebPPL` [@dippl]. The corresponding model code can be found in the associated online repository (file xxxxx). To generate model predictions, we estimated age sensitive parameter distributions for semantic knowledge (by familiar object), speaker informativeness and sensitivity to common ground and then passed them through the model in line with the different ways in which they can be combined and aligned. The resulting predictions come in the form of distributions of developmental trajectories for each object in the congruent and the incongruent condition.

We used the following prior distributions for model parameters. Intercept and slope for sensitivity to common ground: $unif(-2,2)$. For the speaker informativeness we used $unif(-3,3)$ for the intercept $unif(-0,4)$ for the slope. We restricted the slope to be positive because negative values for speaker informativeness are conceptually implausible. For semantic knowledge we used $unif(-3,3)$ for the intercept and $unif(0,2)$ for the slope, again because negative values of semantic knowledge have no meaningful interpretation. For the parameter capturing the variability the object specific trajectories around these overall parameters we used $unif(0,2)$ for the intercept and $unif(0,1)$ for the slope. Some choices regarding these prior distributions were made to ease model convergence. However, please note that all models considered used the same prior distributions.
